noam_data_exploration_v1

### 2024-08-23 noam_data_exploration_v3.Rmd
Added figs and models for diversity synchrony stability relationships

```{r diversity portfolio effects and synchrony quadrat level,  fig.height = 8, fig.width = 10}
(loreau_diversity_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = shannon_mean, y = loreau_synchrony, colour = site, shape = habitat), size = 4, stroke = 1) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

(portfolio_diversity_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = shannon_mean, y = portfolio_effect, colour = site, shape = habitat), size = 4, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

diversity_stability_quadrat$loreau_synchrony_trans <- sv_trans(diversity_stability_quadrat$loreau_synchrony, diversity_stability_quadrat) # for modelling on beta
synchrony_diversity_universal <- glmmTMB(loreau_synchrony_trans ~shannon_mean*site + shannon_mean*habitat + site*habitat + (1|transect/location), family = beta_family(), data = diversity_stability_quadrat, na.action = na.pass)
model.sel(dredge(synchrony_diversity_universal)) # best model as specified
summary(synchrony_diversity_universal)
Anova(synchrony_diversity_universal)
r.squaredGLMM(synchrony_diversity_universal) # Model's distribution-specific variance is negative. Results are not reliable.

>>>>

```{r diversity portfolio effects and synchrony quadrat level,  fig.height = 8, fig.width = 10}
diversity_stability_quadrat$loreau_synchrony_trans <- sv_trans(diversity_stability_quadrat$loreau_synchrony, diversity_stability_quadrat) # for modelling on beta
synchrony_shann_mod <- glmmTMB(loreau_synchrony_trans ~ shannon_mean*habitat*island_side + (1|island_side/site/transect), 
                             family = beta_family(),
                          data = diversity_stability_quadrat, na.action = na.pass)
# model.sel(dredge(synchrony_shann_mod)) # top model is as specified
summary(synchrony_shann_mod)
synchrony_shann_effects <- emtrends(synchrony_shann_mod, pairwise ~ island_side | habitat, var = "shannon_mean")
synchrony_shann_emm <- emmip(synchrony_shann_mod,
                        habitat ~ shannon_mean | island_side,
                        at = list(shannon_mean = seq(0,1.25,0.05)), plotit = F, CIs = T)

filtered_synchrony_div_effects <- synchrony_shann_emm %>%
  left_join(diversity_ranges, by = c("island_side", "habitat")) %>%
  filter(shannon_mean >= min_shannon & shannon_mean <= max_shannon)

(loreau_diversity_plot_quad <-
  ggplot() +
  geom_point(data = diversity_stability_quadrat,
             aes(x = shannon_mean, y = loreau_synchrony, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_synchrony_div_effects, aes(x = shannon_mean, y = inv_logit(yvar), colour = island_side), size = 1) +
  geom_ribbon(data = filtered_synchrony_div_effects, aes(x = shannon_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
                                                         fill = island_side),
             alpha = 0.3) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat) +

@@

synchrony_rich_mod <- glmmTMB(loreau_synchrony_trans ~ richness_mean*habitat*island_side + (1|island_side/site/transect), 
                             family = beta_family(),
                          data = diversity_stability_quadrat, na.action = na.pass)
# model.sel(dredge(synchrony_rich_mod)) # top model is as specified
summary(synchrony_rich_mod)
synchrony_rich_effects <- emtrends(synchrony_rich_mod, pairwise ~ island_side | habitat, var = "richness_mean")
synchrony_rich_emm <- emmip(synchrony_rich_mod,
                        habitat ~ richness_mean | island_side,
                        at = list(richness_mean = seq(0,5,0.1)), plotit = F, CIs = T)

filtered_synchrony_rich_effects <- synchrony_rich_emm %>%
  left_join(diversity_ranges, by = c("island_side", "habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)

(loreau_richness_plot_quad <-
  ggplot() +
  geom_point(data = diversity_stability_quadrat,
             aes(x = richness_mean, y = loreau_synchrony, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_synchrony_rich_effects, aes(x = richness_mean, y = inv_logit(yvar), colour = island_side), size = 1) +
  geom_ribbon(data = filtered_synchrony_rich_effects, aes(x = richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
                                                         fill = island_side),
             alpha = 0.3) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat) +

@@

synchrony_diversity_universal <- glmmTMB(loreau_synchrony_trans ~shannon_mean*site + shannon_mean*habitat + site*habitat + (1|transect/location), family = beta_family(), data = diversity_stability_quadrat, na.action = na.pass)
model.sel(dredge(synchrony_diversity_universal)) # best model as specified
summary(synchrony_diversity_universal)
Anova(synchrony_diversity_universal)
r.squaredGLMM(synchrony_diversity_universal) # Model's distribution-specific variance is negative. Results are not reliable.
